### 2021-09-03 input/proposal.tex
negligible
slowly addressing reviews

At any time, a \process can decide to become a \emph{source}, hence
creating a new partition in the system by executing an exttt{Add}
operation. This partition includes at least its source plus
neighboring \processes that estimate they are closer to this source
than any other one. Such a distance (or \emph{weight}) is
application-dependant: in the context of maintaining distributed
indexes, this would be about link latency.

\begin{definition}[\label{def:partitioning}Partitioning]
  Let $S \subseteq V$ be the set of \underline{s}ources, and $P_s$ be
  the \underline{p}artition including at least \Process~$s$, each
  \process belongs to at most one partition \smash{\small$\forall p,q \in V, \forall
  s_1,s_2 \in S: p \in P_{s_1} \wedge q \in P_{s_2} \implies p \neq q
  \vee s_1 = s_2$}, and there exists at least one path $\pi_{ps}$ of
  \processes that belong to this partition $\forall q \in \pi_{ps}: q
  \in P_s$.
\end{definition}

@@

\begin{definition}[\underline{C}onsistent \underline{p}artitioning (CP)]
  Let $W_{xy} = W_{yx}$ be the positive symmetric \underline{w}eight
  between $x$ and $y$, $\Pi_{xz}$ be the shortest \underline{p}ath
  from $x$ to $z$ the weight of which $|\Pi_{xz}|$ is lower than any
  other path weight, the only consistent partitioning $\mathcal{P}$ is
  a set of partitions $P_{s\in S}$ such that each \process belongs to
  a logical partition comprising its closest source: $\forall p \in
  P_{s_1}: \nexists P_{s_2}$ such that $|\Pi_{s_1p}| > |\Pi_{s_2p}|$.
\end{definition}

@@

\begin{theorem}[\label{theo:bef}\underline{B}est \underline{e}ventual
    \underline{f}orwarding $(BEF) \implies CP$]
%
Assuming reliable communication links where a correct \process $q$
eventually receives the message $m$ \underline{s}ent to it by a
\process $p$ ($s_{pq}(m) \implies r_{q}(m)$), \processes eventually
reach consistent partitioning if each \process eventually forwards its
best known partition.
\end{theorem}

@@

While only adding logical partitions to the distributed system is
straightforward, removing them introduces additional complexity.

@@

\begin{corollary}[\label{theo:dcp}purge + BEF $\implies$
    \underline{D}ynamic \underline{c}onsistent
    \underline{p}artitioning (DCP)]
%
When each \process can \texttt{Add} and \texttt{Del}ete its partition
at any time, consistent partitioning requires
\begin{inparaenum}[(i)]
\item eventual purging of stale messages ($b_p(m) \rightarrow b_p(m')
  \implies \eventually (d_q(m) \implies d_q(m) \rightarrow
  d_q(m''))$);
\item and best eventual forwarding.
\end{inparaenum}
\end{corollary}

>>>>

At any time, a \process can decide to become a \emph{source}, hence
creating a new partition in the system by executing an \texttt{Add}
operation. This partition includes at least its source plus
neighboring \processes that estimate they are closer to this source
than any other one. Such a distance (or \emph{weight}) is
application-dependant: in the context of maintaining distributed
indexes, this would be about link latency that \nodes could monitor
by aggregating \texttt{ping}s.

\begin{definition}[\label{def:partitioning}Partitioning]
  Let $S \subseteq V$ be the set of \underline{s}ources, and $P_s$ be
  the \underline{p}artition including at least \Process~$s$, each
  \process belongs to at most one partition $\forall p,q \in V, \forall
  s_1,s_2 \in S: p \in P_{s_1} \wedge q \in P_{s_2} \implies p \neq q
  \vee s_1 = s_2$, and there exists at least one path $\pi_{ps}$ of
  \processes that belong to this partition $\forall q \in \pi_{ps}: q
  \in P_s$.
\end{definition}

@@

\begin{definition}[\underline{C}onsistent \underline{p}artitioning (CP)]
  Let $W_{xy} = W_{yx}$ be the positive symmetric \underline{w}eight
  between $x$ and $y$, $\Pi_{xz}$ be the shortest \underline{p}ath
  from $x$ to $z$ the weight of which $|\Pi_{xz}|$ is lower than any
  other path weight, with $|\Pi_{xx}|$ being the greatest lower bound
  of $x$, the only consistent partitioning $\mathcal{P}$ is a set of
  partitions $P_{s\in S}$ such that each \process belongs to a logical
  partition comprising its closest source: $\forall p \in P_{s_1}:
  \nexists P_{s_2}$ such that $|\Pi_{s_2p}| < |\Pi_{s_1p}|$.
\end{definition}

@@

\begin{theorem}[\label{theo:bef}\underline{B}est \underline{e}ventual
    \underline{f}orwarding $(BEF) \implies CP$]
%
Assuming reliable communication links where a correct \process $q$
eventually receives the message $m$ \underline{s}ent to it by a
\process $p$ ($s_{pq}(m) \implies r_{q}(m)$), \processes eventually
reach consistent partitioning if each \process eventually
\underline{f}orwards ($f_p(m) \implies \forall \langle p, q\rangle \in
E: s_{pq}(m)$) its best known partition.
\end{theorem}

@@

While only adding logical partitions to the distributed system is
straightforward, removing them introduces additional complexity caused
by concurrent operations.

@@

\begin{theorem}[\label{theo:dcp}purge + BEF $\implies$
    \underline{D}ynamic \underline{c}onsistent
    \underline{p}artitioning (DCP)]
%
When each \process can \texttt{Add} and \texttt{Del}ete its partition
at any time, consistent partitioning requires
\begin{inparaenum}[(i)]
\item eventual purging of stale messages ($b_p(m) \rightarrow b_p(m')
  \implies \eventually (d_q(m) \implies d_q(m) \rightarrow
  d_q(m''))$);
\item and best eventual forwarding.
\end{inparaenum}
\end{theorem}
